---
title: "Lab-2"
author: "Nishanth Girish"
date: "2023-09-29"
output: word_document
---

The objective of this assignment is to solve the following problems using R statistical software using appropriate R commands:

Consider a data of your choice of 50 observations and estimate the following.
1. Select a simple random sample of size 15 units using with and without replacement
procedure and estimate the sample mean along with the standard deviation for both
SRSWOR and SRSWR.
2. Estimate sample mean square for both SRSWOR and SRSWR using the formula.
3. Prove that sample mean is an unbiased estimate of population mean for both SRSWOR
and SRSWR.
4. Prove that sample mean square is an unbiased estimate of population variance in
SRSWR.

DATASET:

We take 50 random observations with mean 5.6 and variance 0.8 and from the normal distribution using the ‘rnorm’ built-in function as shown below:

```{r}
#1.Select a simple random sample of size 15 units using with and without replacement procedure and estimate the sample mean along with the standard deviation for both SRSWOR and SRSWR.

#SRSWOR:
set.seed(77)
x= rnorm(50,5.6,0.8)
x

set.seed(77)
s1 <- sample(x,size = 15) 
s1

mean(s1)

sd(s1)
```


```{r}
#SRSWR:
set.seed(77)
s2 <- sample(x, size = 15, replace = TRUE)
s2

mean(s2)

sd(s2)
```


```{r}
#2.Estimate sample mean square for both SRSWOR and SRSWR using the formula.

#SRSWR 
a1 = s1^2
summ.yisq1 = sum(a1)
b1= 15 * (mean(s1)^2)

sms1 = 1/14 * (summ.yisq1 - b1)  
sms1                              #sample mean square 


#SRSWOR
a2 = s2^2
summ.yisq2 = sum(a2)
b2 = 15* (mean(s2)^2)

sms2 =1/14 * (summ.yisq2 - b2)
sms2                             #sample mean square

```


```{r}
#3.Prove that sample mean is an unbiased estimate of population mean for both SRSWOR and SRSWR.

mean(s1) 
mean(x)        #popln mean will be approximately equal to sample mean                      


mean(s2)
mean(x)       #popln mean will be approximately equal to sample mean 

```


```{r}
#4.Prove that sample mean square is an unbiased estimate of population variance in SRSWR.
popvariance=var(x)    #population variance
popvariance

a1 = s1^2
summ.yisq1 = sum(a1)
b1= 15 * (mean(s1)^2)

sms1 = 1/14 * (summ.yisq1 - b1)  
sms1                               #sample mean square in case of SRSWR

```

