---
title: "Lab-8"
author: "Nishanth Girish"
date: "2023-09-29"
output: word_document
---

Example on stratified random sampling
Consider the mtcars dataset as a population, divide the population
into strata on the basis of type of engine used in cars and draw a random
sample of size 10 from the population by equal allocation and estimate the
average mpg of cars with 95% confidence interval. 

```{r}
library(samplingbook)

```


```{r}
attach(mtcars)
#Since there are two types of engines in the cars so population is divided into two strata i.e. vs=1 and vs=0 and then the sample of size 10 will be allocated to strata using equal allocation : stratasamp()
# Obtaining stratum size and stratum standard deviation (N1, N2, S1, S2)
stratum1=mtcars[mtcars$vs==1, ]
stratum1 #population in stratum 1

N1=sum(stratum1$vs==1)
N1 # stratum1 size
## [1] 14
S1=sqrt(var(stratum1$mpg))
S1 # population standard deviation of stratum 1
```


```{r}
stratum2=mtcars[mtcars$vs==0, ]
stratum2 #population in stratum 2
```


```{r}
mean(stratum1$mpg) # mean of mpg of stratum 1

mean(stratum2$mpg) # mean of mpg of stratum 2

```


```{r}
N2=sum(stratum2$vs==0)
N2 # stratum2 size

S2=sqrt(var(stratum2$mpg))
S2 # population standard deviation of stratum 2
```


```{r}
# The sample of size n=10 has to be drawn using equal allocation using SRSWOR.i.e.
n1=5
n2=5
sample1=stratum1[sample(1:nrow(stratum1), 5, replace=FALSE), ]
sample1 # sample 1 collected from stratum 1
```


```{r}
sample2=stratum2[sample(1:nrow(stratum2), 5, replace=FALSE), ]
sample2 # sample 2 collected from stratum 2
```


```{r}
total_sampled_data=rbind(sample1, sample2)
total_sampled_data # total sample collected using stratified random sampling
```


```{r}
# Estimation of population mean i.e. mean mpg using stratified random sample: stratamean()
nh=as.vector(table(total_sampled_data$vs))
nh

wh=nh/sum(nh)
wh

stratamean(y=total_sampled_data$mpg, h=as.vector(total_sampled_data$vs), wh=wh)

#Extended
stratamean(y=total_sampled_data$mpg, h=as.vector(total_sampled_data$vs), wh=wh, eae=TRUE)
```
Conclusion: the estimated average mgp of cars obtained using equal allocation
of sample of size 10 to strata is 20.86 per gallon with the SE of 1.72 and
the 95% confidence interval is [17.4905,24.2295] means on an average, the
minimum millage of cars can be 17.49 per gallon and the maximum millage of
cars can be 24.23 per gallon with .95 confidence coefficient. Similarly, the
estimated values for each stratum can be interpreted and comparison can also
be done in estimated of each stratum.

