---
title: "Lab-3"
author: "Nishanth Girish"
date: "2023-09-29"
output: word_document
---

The objective of this assignment is to solve the following problems using R statistical software using appropriate R commands:
Take a data of your choice and consider it as population.
1. Draw a random sample of any size which is suitable to you using with replacement procedure and give an estimate for population parameters mean and variance with confidence intervals.  
2. Obtain the sampling distribution of the sample mean and represent it by a histogram.
Comment on its nature.
3. Prove that sample mean is an unbiased estimate of the population mean
4. Verify that variance of the estimate is minimum in SRSWOR than the SRSWR(use formulae to calculate the variances in both cases SRSWOR and SRSWR by taking a sample of the same size and compare the variances) 


```{r}
#1. Drawing a random sample with replacement.

set.seed(123)
x=rnorm(35,5.6,0.8)   
x                      #population


s1=sample(x,size = 10, replace = TRUE)
s1                                        #sample

mean(s1)                     #sample mean

pop_mean= mean(x)           #population mean
pop_mean


a1 = s1^2                   #sample mean square 
summ.yisq = sum(a1)
b1= 10 * (mean(s1)^2)


sms = 1/9 * (summ.yisq - b1)  
sms 

var(x)                      #population variance 

#Confidence interval for the mean.
#alpha = 5%
ll_CI=mean(s1)- 1.96*(sd(s1)/sqrt(length(s1)))  #confidence interval formula.
ll_CI                                           #lower limit

ul_CI=mean(s1)+ 1.96*(sd(s1)/sqrt(length(s1)))
ul_CI                                           #upper limit

CI = c(ll_CI,ul_CI)
CI
```

We observe that the sample mean and population mean are approximately equal and it lies in the confidence interval for mean. The 95% confidence interval for mean is [4.998009, 6.229333].

```{r}
#2. Sampling distribution of the sample mean.


samp_dist <- replicate(10000,mean(sample(x,10, replace = TRUE)))
mean(samp_dist)


hist(samp_dist)
```

We observe that the mean of the sampling distribution is approximately equal to the population mean and sample mean.