---
title: "Lab-7"
author: "Nishanth Girish"
date: "2023-09-29"
output: word_document
---

Take a data of your choice and consider it as the population.
1. Draw a simple random sample of any size which is suitable for you using without replacement (SRSWOR) procedure and give an estimate for population parameters mean, variance and SE.  
2. Estimate the confidence interval for the population mean.
3. Obtain the sampling distribution of the sample mean. Represent the distribution graphically and comment on it.
4. Prove that sample mean is an unbiased estimate of the population mean in SRSWOR.

```{r}
#1.Drawing a simple random sample without replacement and estimating population parameters:
set.seed(77)
x <- rnorm(50,5.6, 0.8)        #population
x
```


```{r}
s1 <- sample(x, size= 10)     #SRSWOR
s1

```


```{r}
mean(s1)              #sample mean is an estimate of population mean

sms<-(1/length(s1))*sum((s1- mean(s1))^2)  #sample mean square is an estimate of population mean square in SRSWOR
sms

pop_var = var(x)           #population variance
pop_var

#std error:
se <- sd(x)/sqrt(length(x))
se

```
Here, we obtained a simple random sample without replacement of size 50 from the normal distribution with mean 5.6 and variance 0.8. We also obtained a few population parameters.

```{r}
#2.Confidence interval for mean. Alpha = 5%
ll_CI <- mean(s1) - 1.96*(sd(s1)/length(s1)) #lower confidence limit
ul_CI <- mean(s1) + 1.96*(sd(s1)/length(s1)) #upper confidence limit

CI <- c(ll_CI,ul_CI)
CI

```
The 95% confidence interval for the population mean is given by [5.2502233, 5.602961].

```{r}
#3.Sampling distribution of the sample mean 
samp_dist <- replicate(10000,mean(sample(x,10, replace = FALSE)))
mean(samp_dist)

hist(samp_dist)      #histogram of the sampling distribution

```

We obtain the sampling distribution of the sample mean and represented it graphically in the form of a histogram. The histogram is a bell-shaped curve.

```{r}
#4.Sample mean is an unbiased estimate of population mean:
mean(s1)        #sample mean

mean(x)         #population mean

```

We observe that the sample mean is approximately equal to the population mean. Hence, we conclude that sample mean is an unbiased estimate of population mean.