---
title: "Sampling Theory Lab-1"
author: "Nishanth Girish"
date: "2023-09-29"
output: word_document
---

The objective of this lab is to solve the following problems using R statistical software using appropriate R commands:
The yield in quintals for paddy crop of 50 villages in a certain area are given below.
1. Select a simple random sample of size 10 units using without replacement procedure and estimate the average yield per sample along with the standard deviation.
2. How many plots can be formed from the total paddy field and find the summary statistics of yield from a single plot.
3. Estimate the average yield of total paddy crop with standard error.

```{r}
# Data:

yield <- c(20,21,32,41,55,22,64,42,28,35,25,25,24,32,75,28,29,38,19,19,16,28
,30,29,29,19,37,34,31,35,29,19,27,42,39,11,27,36,28,24,39,71,44,24 ,34,18,28,10,32,47)
yield

```


```{r}
set.seed(123)

#1.SRSWOR with n=10 from Yield along with mean and SD
s<- sample(yield,10)
s

```


```{r}
mean(s)
sd(s)
```


```{r}
#2.No. of possible samples and summary statistics:
plots<- factorial(50)/((factorial(40)*factorial(10)))
plots

```


```{r}
summary(s)
```

```{r}
#3.Estimating avg yield from yield:
mean(s)  #population mean == sample mean
```


```{r}
se = sd(s)/ sqrt(10)
se

```

