---
title: "Lab-5"
author: "Nishanth Girish"
date: "2023-09-29"
output: word_document
---

Consider a dataset of your choice as a population. Divide the population into a number of strata on the basis of suitable variable and estimate the population mean and population total of variable of interest taking a SRSWOR of appropriate size from stratified population and comment on the results. (Use equal allocation to allocate the sample size)

```{r}
#Importing a dataset from Excel:
library(readxl)

Sample_Data_Set <- read_excel("C:/Users/nisha/Downloads/Sample Data Set.xlsx")
```

The dataset Sample_Data_Set contains details of 682 patients which includes personal details(like Age,Gender,Religion etc.), and some vital health parameters(like BMI,cortisol levels etc.) before and after a treatment was given to them. The dataset also shows whether a patient died or not after the treatment. The variable of interest is the Age column. We are estimating the mean age of patients in the sample.

```{r}
#Loading the package:
library(samplingbook)

mydata <- Sample_Data_Set
mydata           #displaying the data
attach(mydata)  #to avoid using $
```


```{r}
#Consider the dataset Sample_Data_Set and assume that it is the population.
#We divide the population into strata on the basis of whether a patient died or not and draw a random sample of size 10 from the population by equal allocation and estimate the average age of patients with 95% confidence interval.

#Since there are two possibilities in the Death column(0:represents no death & 1:represents death), the population is divided into two strata and then the sample of size 10 will be allocated to strata using equal allocation : stratasamp()

# Obtaining stratum size and stratum standard deviation (N1, N2, S1, S2):

stratum1=mydata[Death==1, ]    #Population in stratum 1
stratum1                       

N1=sum(stratum1$Death==1)      #Stratum 1 size
N1                             

mean(stratum1$Age)             #Mean of Age of stratum 1

S1=sqrt(var(stratum1$Age))     #Population standard deviation of stratum 1
S1                             

```


```{r}
stratum2=mydata[Death==0, ]    #Population in stratum 2
stratum2                       

N2=sum(stratum2$Death==0)      #Stratum 2 size
N2                             

mean(stratum2$Age)             #Mean of Age of stratum 2

S2=sqrt(var(stratum2$Age))     #Population standard deviation of stratum 2 
S2                             

```


```{r}
#The sample of size n=10 has to be drawn using equal allocation using SRSWOR:
n1=5
n2=5

sample1=stratum1[sample(1:nrow(stratum1), 5, replace=FALSE), ]
sample1             #Sample 1 collected from stratum 1

```

```{r}
sample2=stratum2[sample(1:nrow(stratum2), 5, replace=FALSE), ]
sample2             #Sample 2 collected from stratum 2

```

```{r}
total_sampled_data=rbind(sample1, sample2)
total_sampled_data #Total sample collected using stratified random sampling

```


```{r}
#Estimation of population mean(mean age) using stratified random sample: 

nh=as.vector(table(total_sampled_data$Death))
nh                    #Vector of sizes of every stratum

wh=nh/sum(nh)
wh                    #Vector of weights of every stratum

stratamean(y=total_sampled_data$Age, h=as.vector(total_sampled_data$Death),wh=wh)

```

```{r}
#Extensive output for each and every stratum:
stratamean(y=total_sampled_data$Age, h=as.vector(total_sampled_data$Death),
wh=wh, eae=TRUE)

```

Interpretation: The estimated average age of patients obtained using equal allocation of sample of size 10 to strata is 56.9 with the SE of 2.6524 and the 95% confidence interval is [51.7015,62.0985], i.e that means for approximately 95% of the samples of same size coming from the same population with the same procedure of sample selection, the mean age of patient will lie in the interval [51.7015,62.0985]. Similarly, the estimated values for each stratum can be interpreted and comparison can also be done in the estimates of each stratum.